---
title: "Up and running GA4GH API server"
author: "Welliton Souza"
date: '`r Sys.Date()`'
package: "`r pkg_ver('GA4GHclient')`"
output: 
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Up and running GA4GH API server}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8} 
---

# Installing GA4GH API server

This vignette describes how to install an instance of [GA4GH API server](https://github.com/ga4gh/server) for tests and development. This tutorial is meant to be used on Linux-based systems. However, the GA4GHclient package also works on Windows systems.

Instead of installing the required softwares on system we will install locally. This part of the vignette is based on [GA4GH API Demo](http://ga4gh-reference-implementation.readthedocs.io/en/latest/demo.html#demo). For installing system-wide see the [Official Installation Guide](http://ga4gh-reference-implementation.readthedocs.io/en/latest/demo.html#demo).

It is required some system packages. For Ubuntu-based systems. On bash.

```
sudo apt-get install python-dev python-virtualenv zlib1g-dev libxslt1-dev libffi-dev libssl-dev
```

Create an isolated environment for GA4GH API server.

```bash
virtualenv ga4gh-env
source ga4gh-env/bin/activate
```

Install stable version of Ga4GH API server from [Python Package Index](https://pypi.python.org/pypi/ga4gh). This version is recommended for testing GA4GH API server and GA4GHclient package.

```
pip install ga4gh --pre
```

Or install development version from [Official GitHub Repository](https://github.com/ga4gh/server). This version is recommended for developing GA4GHclient package.

```
git clone https://github.com/ga4gh/server.git
cd server
python setup.py install
cd ..
```

Download example data.

```
wget https://github.com/ga4gh/server/releases/download/data/ga4gh-example-data-v4.5.tar
tar xvf ga4gh-example-data-v4.5.tar
```

Run GA4GH API server. It will be available at http://127.0.0.1:8000/.

```
ga4gh_server
```

# Testing GA4GH API server

Now we can test the installation by running GA4GHclient package.

```{r, eval=FALSE}
library(GA4GHclient)
host <- "http://127.0.0.1:8000/"
app(host)
```

For more information about GA4GHclient package.

```{r, eval=FALSE}
vignette("GA4GHclient")
```

# Session Information

```{r}
sessionInfo()
```
